---
title: 'Group Project: T09; DS6'
author: '450132759, 450237777, 470315518, 470396894'
date: "October 5, 2018"
output: 
  prettydoc::html_pretty:
    theme: cayman # prettydoc theme
    highlight: null # syntax highlighting
---
<!-- Please run install.packages("prettydoc") and library(prettydoc) before knitting -->
<style>
@import url('https://fonts.googleapis.com/css?family=Roboto+Mono');
@import url('https://fonts.googleapis.com/css?family=Lato');
body{
  font-family: 'Lato' !important;
  font-size: 12pt;
}

code{
  font-family: 'Roboto Mono' !important;
  font-size: 12px;
}

pre{
  font-family: 'Roboto Mono' !important;
  font-size: 12px
}

td{
  font-family: Lato !important;
  font-size: 12pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center")
```

## Abstract
<!-- A one paragraph summary of what you set out to learn, and what you ended up finding. It should summarise the entire report. -->
The Blacklip Abalone, scientific name Haliotis rubra, is common to several parts of Australia (ref: https://dpipwe.tas.gov.au/sea-fishing-aquaculture/community-resources/fish-facts/abalone-blacklip). They are fished recreationally and also farmed. A data set provided by University of California Irvine, originally from the Department of Primary Industry and Fisheries, Tasmania, consists of several physical characteristics, including the number of rings of the Blacklip Abalone shell. The number of rings acts as proxy for the age of the Abalone. We set out to determine and compare both the best and most practical model of age of the abalone using multiple regression. This could then be used to develop and enforce recreational fishing regulations and also to maintain helathy farmed populations. Beginning with a full model, we employ step backward model selection using the Akaike information criterion (AIC). This is compared with a step forward AIC and also a practical model using easy to measure explantory variables. The practical model first removes highly correlated variables using a test for multicollinearity. We then remove difficult to measure explantory variables and test the performance of the resultant model. Our analysis finds that a practical model performs well compared to the stepwise model and can form the basis for developing best harvesting practices both recreationally and for the Blacklip Abalone Farming industry. 

## Introduction
<!-- A discussion of what questions you are trying to answer. -->
Abalone meat is used as a food delicacy, its shells are a foundation for certain types of jewellery and it functions ecologically to stablize kelp forrests and algae in its rocky reef habitat. Blacklip Abalone reach sexual maturity after three to six years and spawning occurs between Spring and Autumn (ref: https://www.dpi.nsw.gov.au/__data/assets/pdf_file/0009/375858/BlacklipAbalone.pdf). Given the importance of a healthy Abalone population, both in the wild and when farmed, it is crucial to develop, apply and regulate best harvesting practices to maintain healthy Abalone populations. With this in mind, we question how well an unrestriced model using all explanatory variables predicts the age of the Blacklip Abalone. An attempt is then made to build a practical model that can be deployed by farmers and regulators allike that provides a similar level of predictive power but doesn't necessitate killing the Abalone or require the use of specialized equipment. 

## Data Set
<!-- Describe details about how the data set was collected (if known) and the variables in the data set. -->
The data is provided by the Machine Leanring Repository at University of California Irvine. It was orginally obtained from the Department of Primary Industry and Fisheries, Tasmania. The variables are:
\begin{table}[]
\begin{tabular}{llll}
Variable        & Type       & Dimension & Description                 \\
Sex             & nominal    &           & M, F and I (infant)         \\
Length          & continuous & mm        & Longest shell measurement   \\
Diameter        & continuous & mm        & perpendicular to length     \\
Height          & continuous & mm        & with meat in shell          \\
Whole\_weight   & continuous & g         & whole abalone               \\
Shucked\_weight & continuous & g         & weight of meat              \\
Viscera\_weight & continuous & g         & gut weight (after bleeding) \\
Shell\_weight   & continous  & g         & after being dried           \\
Rings           & integer    &           & + 1.5 gives age in years   
\end{tabular}
\end{table}

We treat Rings as the dependent variable, and all others as independent.
The data contains no missing values. Two outliers were noticed for Height and removed. 

### Importing Data

```{r import}
data = read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data", sep = ",")
View(data)
library(tidyverse)
library(gridExtra)
glimpse(data)
```

### Cleaning 
```{r}
#install.packages("reshape")
library(reshape)
library(ggplot2)
new_names <- c("Sex", "Length", "Diameter", "Height", "Whole_weight", "Shucked_weight", "Viscera_weight", "Shell_weight", "Rings")

colnames(data) <- new_names

summary(data)

#Length, Diameter, Height in mm
#Weight in g

#Outliers
melt(data) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(outlier.size=2, outlier.colour='blue') + facet_wrap(~ variable, scales = 
  "free") 

#Remove Height = 0.515mm and 1.13mm as it is greater than Q3 + (1.5*IQR) = 0.24
data <- subset(data, data$Height < 0.5)
ggplot(data, aes(y = Height)) +
geom_boxplot(outlier.size=3, outlier.colour='blue')

#Remove measurements that equals to 0 as well as weight s.t. `Shucked weight` + `Viscera weight` + `Shell weight` > `Whole weight`
data <- data %>% filter(Length, Diameter, Height, `Whole_weight`, `Shucked_weight`, `Viscera_weight`, `Shell_weight` != 0) 

#Combine Male and Female
data <- data %>% mutate(Sex = case_when(Sex == "F" ~ "NI", Sex == "M" ~ "NI", Sex == "I" ~ "I"))

#Multivariate Normality
#install.packages("energy")
#install.packages("mvtnorm")
#library(energy)
#library(mvtnorm)
#data_less_sex <- data %>% select(-Sex)
#mvnorm.e(data_less_sex)

#install.packages("MVN")
#library(MVN)
### NO LONGER NEEDED AS NOT REGRESSIN FOR LIN###
#result <- mvn(data = data_less_sex, mvnTest = "dh")
#result$multivariateNormality
#result <- mvn(data = data_less_sex, mvnTest = "dh", univariatePlot = "histogram")

#Check weights
data %>%
  mutate(sum_weight = Shell_weight + Viscera_weight + Shucked_weight) %>%
  summarise(bad = sum(sum_weight > Whole_weight)) 


```

After assigning new variable names to each column, we checked the assumption for multicollinearity. Linear regression assumes that there is little or no multicollinearity in the data; however, from the ggpairs output, most of the variables were relatively highly correlated with one another. The variables with few of the highest correlations were between length and diameter, as well as whole weight and the other weight predictors. Moreover, we found out that despite the difference in their means, male and female seemed to have similar distributions. So from such results, we decided to combine male and female into one category along with removing either length, or diameter from the model.

## Analysis
<!-- Describe how you used multiple regression to analyse the data set. Specifically, you should discuss how you carried out the steps in analysis discussed in class, i.e., exploration of data to find an initial reasonable model, checking the model and changes to the model based on your checking of the model. -->
```{r}
# basic testing - to get a feel for things - not necessarily for preso or report

# some stats
data %>%
  group_by(Sex) %>%
  summarize(mean_wgt = mean(Whole_weight), mean_lenght = mean(Length), mean_rings = mean(Rings), min_rings = min(Rings), max_rings = max(Rings))

# correlation pairs:
library(GGally)
# pairs = ggpairs(data, aes(colour = Sex, alpha = 0.8), title="Pairs plot for abalone dataset") + theme_minimal(base_size = 8)
# ggsave("report/ggpairs.png", pairs, dpi=600)
#ggpairs(data)   # function takes a long time to execute.

# how many of each sex, m/f/infant
data %>%
  group_by(Sex) %>%
  summarize(n = n())


# Full model 
full_model <- lm(Rings ~ ., data = data)
summary(full_model)
t(round(broom::glance(full_model), 2))


```

## Actual Analysis:
The first step is to determine an appropriate model for multiple regression of the data. We use the AIC model selection method in the backwards direction to achieve this. We begin with the full model and remove the least informative variable after every iteration. Crucially, we undergo a log transformation of the dependent variable Rings to ensure normality and homoscedasticity assumptions are held for regression. 

```{r}
# Transform the dependent variable 
data = data %>% mutate(Rings = log(Rings))

# Initialise the full model under M0:
M1 = lm(Rings~.,data=data)

# Use backwards selection of AIC to determine which independent variables 
# would have a linear relationship with the log(Rings) variable.
step.back.aic = step(M1,direction = "backward",trace=FALSE)
summary(step.back.aic)
summary(step.back.aic)$r.squared

# Output our results.
drop1(step.back.aic,test = "F")
library(kableExtra)
data.frame(step.back.aic$coefficients) %>% kable(caption = "Coefficients of Model:") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed"),position="center",full_width = F)

# Check if our assumptions for regression are valid.
library(ggfortify)
autoplot(step.back.aic,which = 1:2)
```

The model determined that the following independent variables Sex, Diameter, Height, Shucked_weight and Viscera_weight all have a relationship with the log(Rings). Their respective p-values are smaller than the critical value and thus we reject the null hypothesis of no relationship with the log(Rings) variable. Interestingly enough, the degrees of freedom for sex is 2, meaning infants have a more significant effect on the number of rings than males with respect to females. Looking at the qqnorm plot, most of the points are close to the 45 degree line. Hence the normality assumption holds. Furthermore, the residual points are symmetrically distributed above and below the zero line and their spread is fairly constant. Hence the linearity and homoscedasticity assumptions hold as well. Our r-square value is 0.57 which is good as 57% of the log(Rings) variables can be explained by our constructed model.

However, the variables chosen by the model, including Sex, Shucked_weight and Viscera_weight can only be determined after cracking open the abalone. This defeats the purpose of the model, as we want to predict as accurate as possible the age of the abalone without opening it to ensure sustainable and efficient farming practices. Thus, a more practical model is chosen with two parameters which are easy to measure without damaging the abalone: weight and diameter. 

```{r}
M2 = lm(Rings~Whole_weight+Diameter,data=data)
summary(M2)
library(ggfortify)
autoplot(M2,which=1:2)
```

As we can see, the r2 value is much higher (0.69 vs 0.57). This suggests that the practical model is a much better fit of the data than the model determined by the AIC method. In addition, it seems as if the normality, linearity and homoscedasticity assumptions all hold as well from looking at the residual and the qqplot. 

## K-folds Cross Validation

```{r}
library(caret)
cv_full = train(
  Rings ~ Sex + Length + Diameter + Height + Whole_weight + Shucked_weight + 
    Viscera_weight + Shell_weight,
  data, 
  method = "lm", 
  trControl = trainControl(
    method = "cv", number = 10, verboseIter = FALSE
    )
)
cv_full

cv_simple = train(
  Rings ~ Sex + Diameter + Height + Whole_weight + Shucked_weight + 
    Viscera_weight + Shell_weight, 
  data, 
  method = "lm", 
  trControl = trainControl(
    method = "cv", number = 10, verboseIter = FALSE
    )
)
cv_simple

ggpairs(data, aes(colour = Sex, alpha = 0.8), title="Pairs plot for abalone dataset") + theme_grey(base_size = 8)


```


## Infant Model
```{r}
data.infant = data
data.infant = data.infant %>% mutate(Sex = case_when(Sex == "F" ~ "NI", Sex == "M" ~ "NI", Sex == "I" ~ "I"))

cv_full = train(
  Rings ~ Sex + Length + Diameter + Height + Whole_weight + Shucked_weight + 
    Viscera_weight + Shell_weight,
  data.infant, 
  method = "lm", 
  trControl = trainControl(
    method = "cv", number = 10, verboseIter = FALSE
    )
)
cv_full

cv_simple = train(
  Rings ~ Sex + Diameter + Height + Whole_weight + Shucked_weight + 
    Viscera_weight + Shell_weight, 
  data.infant, 
  method = "lm", 
  trControl = trainControl(
    method = "cv", number = 10, verboseIter = FALSE
    )
)
cv_simple
```
In the infant model, we combined male and female abalone into the category NI(non-infant), while keeping infants as I.

We observe a slightly higher p-value.

The full model and the simple model both give around the same RMSE and MAE, which indicates that simplifying our model did not negatively impact the accuracy of our prediction.

## Exhaustive Search
```{r}
library(leaps)
regsubset.out = regsubsets(Rings ~ . , data = data, method = "exhaustive")
regsubset.summary = summary(regsubset.out)
plot(regsubset.summary$rsq, xlab = "Number of Variables", ylab = "R Square", type = "b")
plot(regsubset.summary$cp, xlab = "Number of Variables", ylab = "Mallow's Cp", type = "b")
plot(regsubset.out, scale = "Cp")
```
Maybe combining sex into infant/noninfant would be better model? And can also look at removing Length from the model.

## Taking a more appropriate model
```{r}
M1 = lm(Rings~Whole_weight+Diameter+Height,data=data)
summary(M1)
autoplot(M1,which=1:2)
```
## Results
<!-- Provide inferences about the questions of interest. -->


## Discussion and Conclusion
<!-- Describe any limitations of your analysis and how they might be overcome in future research and provide brief conclusions about the results of your study. -->

## References

Warwick J Nash, Tracy L Sellers, Simon R Talbot, Andrew J Cawthorn and Wes B Ford (1994) "The Population Biology of Abalone (_Haliotis_ species) in Tasmania. I. Blacklip Abalone (_H. rubra_) from the North Coast and Islands of Bass Strait", Sea Fisheries Division, Technical Report No. 48 (ISSN 1034-3288), accessed via URL:
https://archive.ics.uci.edu/ml/datasets/abalone

