---
title: 'Modelling the age of abalone: T09; DS6'

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
author: '450132759, 450237777, 470315518, 470396894'


abstract: |
  Write abstract here:



keywords:
  - one
  - two
 

## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle

bibliography: pnas-sample.bib
csl: pnas.csl

## change to true to add optional line numbering
lineno: false

output: rticles::pnas_article
---


Introduction {#introduction .unnumbered}
-------------------

A discussion of what questions you are trying to answer.



Data Set {#data-set .unnumbered}
----------------------



Describe details about how the data set was collected (if known) and the variables in the data set.

```{r import, echo=FALSE, message=FALSE}
data = read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data", sep = ",")
View(data)
library(tidyverse)
library(gridExtra)
##glimpse(data)
```

```{r echo=FALSE, message=FALSE, results="hide"}
#install.packages("reshape")
library(reshape)
library(ggplot2)
new_names <- c("Sex", "Length", "Diameter", "Height", "Whole_weight", "Shucked_weight", "Viscera_weight", "Shell_weight", "Rings")

colnames(data) <- new_names

#summary(data)

#Length, Diameter, Height in mm
#Weight in g

#Outliers
outliers <- melt(data) %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot(outlier.size=2, outlier.colour='blue') + 
 facet_wrap(~ variable, scales ="free") 

#Remove Height = 0.515mm and 1.13mm as it is greater than Q3 + (1.5*IQR) = 0.24
data <- subset(data, data$Height < 0.5)
plot1 <- ggplot(data, aes(y = Height)) + geom_boxplot(outlier.size=3, outlier.colour='blue')

#Remove measurements that equals to 0 as well as weight s.t. `Shucked weight` + `Viscera weight` + `Shell weight` > `Whole weight`
data %>% filter(Length, Diameter, Height, `Whole_weight`, `Shucked_weight`, `Viscera_weight`, `Shell_weight` == 0) %>% filter(`Shucked_weight` + `Viscera_weight` + `Shell_weight` > `Whole_weight`)
summary(data)

#Multivariate Normality
#install.packages("energy")
#install.packages("mvtnorm")
library(energy)
library(mvtnorm)
data_less_sex <- data %>% select(-Sex)
#mvnorm.e(data_less_sex)

#install.packages("MVN")
library(MVN)
#result <- mvn(data = data_less_sex, mvnTest = "dh")
#result$multivariateNormality
#result <- mvn(data = data_less_sex, mvnTest = "dh", univariatePlot = "histogram")

# check weights
data %>%
  mutate(sum_weight = Shell_weight + Viscera_weight + Shucked_weight) %>%
  summarise(bad = sum(sum_weight > Whole_weight)) 
```


Analysis {#analysis .unnumbered}
-------------------

Describe how you used multiple regression to analyse the data set. Specifically, you should discuss how you carried out the steps in analysis discussed in class, i.e., exploration of data to find an initial reasonable model, checking the model and changes to the model based on your checking of the model.


```{r plot2, echo=FALSE, out.width='0.98\\columnwidth', fig.cap="Outliers test", }
plot1

```



Results {#results .unnumbered}
-----------------

Provide inferences about the questions of interest.



Discusion {#discussion .unumbered}
-----------------

Discussion stuff here

```{r warning=FALSE, message=FALSE, echo=FALSE}
library(kableExtra)
kable(data)
```


Conclusion {#conclusion .unumbered}
-----------------
```{r plot3, echo=FALSE, out.width='0.98\\columnwidth', fig.cap="Outliers test", }
plot1

```


References {#references .unnumbered}
----------

References should be cited in numerical order as they appear in text;
this will be done automatically via bibtex, e.g. @belkin2002using and
@berard1994embedding [@coifman2005geometric]. All references, including
for the SI, should be included in the main manuscript file. References
appearing in both sections should not be duplicated. SI references
included in tables should be included with the main reference section.

Data Archival {#data-archival .unnumbered}
-------------

PNAS must be able to archive the data essential to a published article.
Where such archiving is not possible, deposition of data in public
databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata,
and others outlined in the Information for Authors, is acceptable.



Language-Editing Services {#language-editing-services .unnumbered}
-------------------------

Prior to submission, authors who believe their manuscripts would benefit
from professional editing are encouraged to use a language-editing
service (see list at www.pnas.org/site/authors/language-editing.xhtml).
PNAS does not take responsibility for or endorse these services, and
their use has no bearing on acceptance of a manuscript for publication.

![Placeholder image of a frog with a long example caption to show
justification setting.<span data-label="fig:frog"></span>](frog.png)

Digital Figures {#sec:figures .unnumbered}
---------------

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for
figures that will appear in the main text, and images must be final
size. Authors may submit U3D or PRC files for 3D images; these must be
accompanied by 2D representations in TIFF, EPS, or high-resolution PDF
format. Color images must be in RGB (red, green, blue) mode. Include the
font files for any text.

Figures and Tables should be labelled and referenced in the standard way
using the `\label{}` and `\ref{}` commands.

Figure \[fig:frog\] shows an example of how to insert a column-wide
figure. To insert a figure wider than one column, please use the
`\begin{figure*}...\end{figure*}` environment. Figures wider than one
column should be sized to 11.4 cm or 17.8 cm wide.

Single column equations {#single-column-equations .unnumbered}
-----------------------

Authors may use 1- or 2-column equations in their article, according to
their preference.

To allow an equation to span both columns, options are to use the
`\begin{figure*}...\end{figure*}` environment mentioned above for
figures, or to use the `\begin{widetext}...\end{widetext}` environment
as shown in equation \[eqn:example\] below.

Please note that this option may run into problems with floats and
footnotes, as mentioned in the [cuted package
documentation](http://texdoc.net/pkg/cuted). In the case of problems
with footnotes, it may be possible to correct the situation using
commands `\footnotemark` and `\footnotetext`.

$$\begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{aligned}$$



<!-- pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->

Supporting Information (SI) {#supporting-information-si .unnumbered}
---------------------------


The main text of the paper must stand on its own without the SI. Refer
to SI in the manuscript at an appropriate point in the text. Number
supporting figures and tables starting with S1, S2, etc. Authors are
limited to no more than 10 SI files, not including movie files. Authors
who place detailed materials and methods in SI must provide sufficient
detail in the main text methods to enable a reader to follow the logic
of the procedures and results and also must reference the online
methods. If a paper is fundamentally a study of a new method or
technique, then the methods must be described completely in the main
text. Because PNAS edits SI and composes it into a single PDF, authors
must provide the following file formats only.

### SI Text {#si-text .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference).

### SI Figures {#si-figures .unnumbered}

Provide a brief legend for each supporting figure after the supporting
text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or
GIF format; figures may not be embedded in manuscript text. When saving
TIFF files, use only LZW compression; do not use JPEG compression. Do
not save figure numbers, legends, or author names as part of the image.
Composite figures must be pre-assembled.

### 3D Figures {#d-figures .unnumbered}

Supply a composable U3D or PRC file so that it may be edited and
composed. Authors may submit a PDF file but please note it will be
published in raw format and will not be edited or composed.

### SI Tables {#si-tables .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a
PDF with the same file name for visual reference); include only one
table per file. Do not use tabs or spaces to separate columns in Word
tables.

### SI Datasets {#si-datasets .unnumbered}

Supply Excel (.xls), RTF, or PDF files. This file type will be published
in raw format and will not be edited or composed.

### SI Movies {#si-movies .unnumbered}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media
(wmv), animated GIF (gif), or MPEG files and submit a brief legend for
each movie in a Word or RTF file. All movies should be submitted at the
desired reproduction size and length. Movies should be no more than 10
MB in size.

### Still images {#still-images .unnumbered}

Authors must provide a still image from each video file. Supply TIFF,
EPS, high-resolution PDF, JPEG, or GIF files.

### Appendices {#appendices .unnumbered}

PNAS prefers that authors submit individual source files to ensure
readability. If this is not possible, supply a single PDF file that
contains all of the SI associated with the paper. This file type will be
published in raw format and will not be edited or composed.


<!-- Leave these lines as they are at the end of your .Rmd file to ensure placement of methods & acknowledgements sections before the references-->
\showmatmethods
\showacknow
\pnasbreak
